Food-web complexity flattens the fitness landscape of an insect herbivore

### 2018-06-22 elsevier_test/elsevier_test.tex
updating methods section

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short Paper},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{Short Paper}
    \author[Some Institute of Technology]{Alice Anonymous\corref{c1}}
   \ead{alice@example.com} 
   \cortext[c1]{Corresponding Author}
    \author[Another University]{Bob Security}
   \ead{bob@example.com}

@@

  \begin{abstract}
  Studies of natural selection and fitness landscapes usually treat the
  network of interacting species as a ``black box''. Given that the loss
  of biodiversity is simplifying the structure of ecological networks,
  there is a pressing need to answer the question: how does network
  complexity affect natural selection and the fitness landscape of
  associated species? To answer this question, we conducted a field
  experiment that manipulated the complexity of a food web associated with
  a galling insect herbivore. To maintain complex food webs, we allowed
  the entire community of natural enemies to attack insect galls on 64
  plants in a common garden setting. To create simple food webs, we
  excluded a guild of three larval parasitoids by bagging galls on 64
  different plants; therefore, mortality in this treatment was primarily
  due to a single egg parasitoid that attacks prior to gall formation. We
  then measured herbivore survival as a function of three key gall traits
  in each treatment. We found that more traits were under selection in the
  simple vs.~complex food web. This occurred because different parasitoid
  species impose different selection pressures on gall traits, thereby
  minimizing relative fitness differences among insect galls with
  different phenotypes. Our work suggests that more complex food webs
  allow phenotypic variation to persist, which could facilitate subsequent
  adaptive evolution to environmental change.
  \end{abstract}

@@

Biological diversity -- from genes, to phenotypes, to species -- has
fascinated evolutionary biologists for decades. However, the explicit
context (i.e.~ecological environment) is often abstracted and not
examined in detail in studies of natural selection. Thus, we have a poor
understanding of how biological diversity itself, imposes natural
selection, and the potential feedbacks that may emerge.

@@

We sought to bridge this gap through a field experiment that examines
how food-web complexity alters the fitness landscape of species embedded
within this food web.

@@

To isolate the effects of coastal willow (\textit{S. hookeriana} Barratt
ex Hooker) genetic variation on the plant-insect food web, we used a
common garden experiment consisting of 26 different willow genotypes (13
males; 13 females), located at Humboldt Bay National Wildlife Refuge
(HBNWR) (40°'53``N, 124°12'4''W) near Loleta, California, USA. Willow
genotypes were collected from a single population of willows growing
around Humboldt Bay. While relatedness among these genotypes is unknown,
their phenotypes in multivariate trait space are quite distinct from
each other ({\textbf{???}}), suggesting that we can treat them as
independent from one another. This common garden was planted in February
2009 with 25 clonal replicates (i.e.~stem cuttings) of each willow
genotype in a completely randomized design in two hectares of a former
cattle pasture at HBNWR. Willows in our garden begin flowering in
February and reach their peak growth in early August. During this study,
willows had reached 2 - 4 m in height. Further details on the genotyping
and planting of the common garden are available in (Barbour et al.
2015).

We conducted our study within a common garden consisting of 26 different
willow genotypes (13 males; 13 females), located at Humboldt Bay
National Wildlife Refuge (HBNWR) (40°40'53``N, 124°12'4''W) near Loleta,
California, USA. Willow genotypes were collected from a single
population of willows growing around Humboldt Bay. While relatedness
among these genotypes is unknown, their phenotypes in multivariate trait
space are quite distinct from each other (Barbour et al. 2016),
suggesting that we can treat them as independent from one another. This
common garden was planted in February 2009 with 25 clonal replicates
(i.e.~stem cuttings) of each willow genotype in a completely randomized
design in two hectares of a former cattle pasture at HBNWR. Willows in
our garden begin flowering in February and reach their peak growth in
early August. During this study, willows had reached 2 - 4 m in height.
Further details on the genotyping and planting of the common garden are
available in (Barbour et al. 2015). We conducted this experiment across
8 different plant genotypes that span the range of trait variation
(Barbour et al. 2015).

We setup our food-web manipulation soon after galls began developing on
\textit{S. hookeriana} in early June of 2013. On treatment plants, we
enclosed 14 galled leaves with organza bags (MANUFACTORER DETAILS) to
exclude three parasitoid species that attack during larva development
(hereafter larval parasitoids). This treatment did not exclude the egg
parasitoid \textit{Platygaster} sp. which attacks prior to gall
initiation (note that in Cecidomyiid midges, larva initiate gall
development CITE). On control plants, we used flagging tape to mark 14
galled leaves per plant, allowing the full suite of parasitoids to
attack \textit{Iteomyia}. Marking galls with flagging tape ensured that
we compared control and treatment galls with similar phenology when we
collected galls later in the season. In late August 2013, we collected
marked and bagged galls from each plant. We placed galls in 30 mL
transport vials and allowed them to complete development for 4 months at
room temperature in the lab. We opened galls under a dissecting scope
and determined whether larva survived to pupation, our measure of
fitness, or were parasitized.

We collected data on three different phenotypes for each larva that have
been shown to be important in our work (cite Barbour et al. 2016) and in
other work with Cecidomyiid midges (CITE). First, we measured the size
of each gall chamber to the nearest 0.01 mm at its maximum diameter
(perpendicular to the direction of plant tissue growth). Second, we
counted the number of chambers in each gall, which is indicative of the
number of larva per gall. All larva collected from the same gall were
scored with the same `number of larva per gall' phenotype. This trait is
indicative of the clutch size of adult females. \%This trait likely
reflects a combination of both larva feeding behavior as well as adult
female preferences and clutch size (CITE Art Weis' work). Third, we
estimated gall density as the number of larva per 100 shoots. We did
this by counting the number of gall chambers on five randomly sampled
branches per tree. To account for potential differences in the number of
shoots per branch for each plant genotype (CITE other willow work), we
then counted the number of shoots on the fifth branch to estimate the
number of larva per 100 shoots for each plant. All larva collected from
the same plant were scored with the same gall density phenotype. This
phenotype is indicative of the preference of female midges for
particular plant traits (hereafter `female preference').

@@

To account for the correlated structure of our gall phenotypes (female
preference at plant-level; clutch size at gall-level; chamber size at
chamber-level), we specified gall ID nested within plant ID nested
within plant genotype as random intercepts in our statistical models.

@@

We used a spline-based semiparametric regression (Schluter 1988,
Morrissey and Sakedra 2014). These analyses are desirable because they
lead to inferences of the form of selection that make few
\textit{a priori} assumptions (Schluter 1988), and can now be used to
quantify standardized selection gradients (Morrissey and Sakrera 2014).

We then calculated selection gradients as the partial derivatives in
absolute fitness (larva survival) with response to multivariate
phenotype using the \textit{gsg} package in R (cite Morrissey and R
project). We standardized phenotypes (mean=0, SD=1), so that first-order
derivatives correspond to the intensity of directional selection
(\(\beta\)\textsubscript{trait}), while second-order derivatives
correspond to intensity of nonlinear (\(\gamma\)\textsubscript{trait})
and correlational selection
(\(\gamma\)\textsubscript{trait$_i$,}\textsubscript{trait$_j$}) and are
thus comparable within this study as well as to others.

Our GLMMs are useful for testing the effects of food-web complexity on
the fitness landscape within the context of our experimental design;
however, coefficients from GLMMs cannot be easily converted into
quantitative estimates of selection gradients (cite Morrissey's work).
Therefore, for each treatment, we fit separate generalized additive
models (GAMs) for gall traits that we identified as being under
selection from our GLMMs. Estimating selection gradients from GAMs can
give insight to both linear and non-linear selection gradients, which we
assumed were all linear for our GLMMs, given the complexity of already
fitting up to 4-way interactions in these models. For the number of
larva per gall and gall density, we aggregated larva survival at the
gall or plant level, respectively, to avoid pseudoreplication in our
GAMs.

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Food-web complexity flattens the fitness landscape of an insect herbivore},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

  \title{Food-web complexity flattens the fitness landscape of an insect
herbivore}
    \author[Department of Evolutionary Biology and Environmental Studies, University
of Zurich, Zurich, Switzerland]{Matthew A. Barbour\corref{c1}}
   \ead{matthew.barbour@ieu.uzh.ch} 
   \cortext[c1]{Corresponding Author}
    \author[Another University]{Jordi Bascompte}
   \ead{jordi.bascompte@ieu.uzh.ch}

@@

  \begin{abstract}
  true
  \end{abstract}

@@

Biological diversity -- from genes, to phenotypes, to species -- has
fascinated evolutionary biologists for decades.

Much of this biological diversity has been shaped by natural selection
via trophic interactions, such as resource competition (cite charater
displacement), mutualistic exchanges of resource (cite? ), and predation
(cite recent Heath and Stireman paper).

NEED to recognize that evolutionary biologists have begun to explore how
community context affects evolutionary change (work by Sharon Strauss,
Casey terHorst, Lutz Becks, etc.). These results have begun to show
interesting patterns whereby the composition of species in a community
can alter the direction and strength of natural selection imposed on
species embedded within these communities (cite).

Put another way, these results have begun to show biological diversity,
in terms of differences between species, can drive evolutionary change.

While there is clear evidence for pairwise trophic interactions to drive
evolution by natural selection, its unclear how biological diversity
itself imposes natural selection and drives evolutionary change.

Exploring the effects of biological diversity requires an explicit
examination of the network structure of trophic interactions between
species in a community. Theoretical models have begun to examine how the
network structure of species interactions drives evolutionary change
(Nuismer paper; Guimeras paper; Ecology Letters paper from a spanish
guy\ldots{}); however, we are currently lacking experimental tests.

@@

Food-web complexity may influence selection gradients in at least two
ways. First, if more diverse predator communities are more efficient as
suppressing prey densities (e.g.~biodiversity-ecosystem fucntion; Ives
2005 Ecology Letters), then this will result in lower mean fitness. A
reduction in mean fitness, all else equal, will intensify natural
selection and thus could increase the rate of evolutionary change.
Alternatively, if predators are functionally distinct, more diverse
communities can reduce the strength of selection. This is because each
predator has a different functional relationship between prey phenotype
and the probability of an interaction.

Here, we conducted a field experiment to test the effect of food-web
complexity on the fitness landscape of a species embedded within this
food web. To do this, we used a common garden experiment with a host
plant (\emph{Salix hooeriana}), an abundant herbivore (\emph{Iteomyia
salicisverruca}), and the diverse community of insect parasitoids that
attack it. Prior work in this system has shown that there is directional
selection for larger galls, likely because larger galls provide more of
a refuge from parasitoid attack. However, there is also evidence that
different parasitoid species impose differential selection on gall
phenotypes.

@@

We conducted our study within a four-year old common garden of
\emph{Salix hookeriana} located at Humboldt Bay National Wildlife Refuge
(HBNWR) (40°40'53``N, 124°12'4''W) near Loleta, California, USA. This
common garden consists of 26 different willow genotypes that were
collected from a single population of willows growing around Humboldt
Bay. Clonal replicates (stem cuttings) of each genotype were planted in
a completely randomized design in two hectares of a former cattle
pasture at HBNWR. Willows in our garden begin flowering in February and
reach their peak growth in early August. During this study, willows had
reached 5 - 9m in height. Further details on the genotyping and planting
of the common garden are available in (Barbour et al. (2015)).

We setup our food-web manipulation soon after galls began developing on
\emph{S. hookeriana} in early June of 2013. We manipulated gall exposure
to parasitoids across 128 plants from 8 different plant genotypes that
span the range of trait variation observed in this willow population
(Barbour et al. (2015)). On treatment plants (8 replicates per
genotype), we enclosed 14 galled leaves with organza bags (MANUFACTORER
DETAILS) to exclude three parasitoid species that attack during larva
development (hereafter larval parasitoids). This treatment did not
exclude the egg parasitoid \emph{Platygaster} sp. which attacks prior to
gall initiation (note that in Cecidomyiid midges, larva initiate gall
development CITE). On control plants, we used flagging tape to mark 14
galled leaves per plant, allowing the full suite of parasitoids to
attack \emph{Iteomyia}. Marking galls with flagging tape ensured that we
compared control and treatment galls with similar phenology when we
collected galls later in the season. In late August, we collected marked
and bagged galls from each plant. We placed galls into 30 mL vials and
kept them in the lab for 4 months at room temperature. We then opened
galls under a dissecting scope and determined whether larva survived to
pupation (our measure of fitness) or were parasitized.

We collected data on three different phenotypes for each larva that have
been shown to be important in our work (Barbour et al. (2016)) and in
other work with Cecidomyiid midges (Weis, Price, and Lynch (1983),
Heath, Abbot, and Stireman (2018)). First, we measured the size of each
gall chamber to the nearest 0.01 mm at its maximum diameter
(perpendicular to the direction of plant tissue growth). Our previous
work has shown that a larger gall diameter provides a refuge for larva
from parasitoid attack (Barbour et al. (2016)). Second, we counted the
number of chambers in each gall, which is indicative of the clutch size
of adult female midges (Weis, Price, and Lynch (1983)). All larva
collected from the same multi-chambered gall were scored with the same
clutch size. Third, we estimated gall density as the number of chambers
per 100 shoots per plant. We did this by summing the number of chambers
across all galls found on five randomly sampled branches per tree. To
account for potential differences in the number of shoots per branch for
each plant genotype (CITE other willow work), we then counted the number
of shoots on the fifth branch to estimate the number of larva per 100
shoots for each plant. All larva collected from the same plant were
scored with the same gall density phenotype. This phenotype is
indicative of the preference of female midges for particular plant
traits (hereafter `female preference'). The measurement of insect
densities on a plant is a commonly used index for measuring oviposition
preference (Gripenberg et al. (2010)), although caution must be taken in
inferring `preference' (Singer (1986)). This is because oviposition is a
complex process and the number of individuals observed on plants in the
field can be influenced by processes other than preference. While we
recognize this, since our data comes from a randomized placement of
genotypes and that we can still detect the presence of an individual
even if it dies before development (empty chamber lacking an exit hole
and without any evidence of larva or parasitoid). We cannot control for
egg predation, although this likely plays a comparatively minor role in
determining gall survival (Hawkins, Cornell, and Hochberg (1997)).

To identify the appropriate level of model complexity for testing the
effects of food-web complexity on the fitness landscapes, we compared
models using Aikaike Information Criteria. The maximal complexity we
explored was a generalized additive mixed model that fit cubic splines
to each trait as well as linear interactions between a maximum of two
traits as well as an interaction with food-web treatment. We then
examined simpler models where

This analysis enabled us to explore the appropriate complexity of the
model to include for our analyses (i.e.~include non-linear and
correlational selection gradients and whether they varied with food-web
treatment).

@@

To quantify selection gradients, we fit separate statistical models to
data from each food-web treatment. We used the method of Frederic J
Janzen and Hal S Stearn (1998) to calculate selection gradients and used
parametric bootstrapping to calculate their 95\% confidence intervals
(Bolker et al. (2009)).

To account for the correlated structure of our gall phenotypes
(oviposition preference at plant-level; clutch size at gall-level; gall
diameter at chamber-level), we specified gall ID nested within plant ID
nested within plant genotype as random intercepts in our statistical
models.

@@

\section{Results}\label{results}

We found that more phenotypic traits were under selection in the simple
vs.~complex food web. In both complex and simple food webs, gall
diameter was under strong directional selection, with larger galls
resulting in higher larval survival (complex Beta = ; simple Beta =
)(Fig. 2A). In complex food webs, there was no evidence of selection on
clutch size (\(\beta_{clutch}=\)) or female preference
(\(\beta_{preference}=\))(orange lines in Fig. 2B,C). In simple food
webs, however, clutch size and female preference were under strong
directional selection, with smaller clutch sizes and weaker preferences
resulting in higher larval survival (blue lines in Fig. 2B,C). These
different selection pressures resulted in different adaptive landscapes
in complex vs.~simple food webs, with evidence for more rugged
landscapes in the simple rather than the complex food web (Fig. 3).
Depending on the trait combinations used to create the landscape, we
found that the ruggedness of the adaptive landscape ranged from 10\%
higher (Fig. 3A) to 274\% higher (Fig. 3C) in the simple vs.~complex
food web. Our model comparison suggested that it was unnecessary to test
for the effects of non-linear or correlational selection gradients (ref.
supp. mat.).

\section{Discussion}\label{discussion}

@@

\hypertarget{refs}{}
\hypertarget{ref-Barbour2016}{}
Barbour, Matthew A, Miguel A Fortuna, Jordi Bascompte, Joshua R
Nicholson, Riitta Julkunen-Tiitto, Erik S Jules, and Gregory M
Crutsinger. 2016. ``Genetic Specificity of a Plant--insect Food Web:
Implications for Linking Genetic Variation to Network Complexity.''
\emph{Proceedings of the National Academy of Sciences} 113 (8). National
Acad Sciences: 2128--33.

\hypertarget{ref-Barbour2015}{}
Barbour, Matthew A, Mariano A Rodriguez-Cabal, Elizabeth T Wu, Riitta
Julkunen-Tiitto, Carol E Ritland, Allyson E Miscampbell, Erik S Jules,
and Gregory M Crutsinger. 2015. ``Multiple Plant Traits Shape the
Genetic Basis of Herbivore Community Assembly.'' \emph{Funct. Ecol.} 29
(8): 995--1006.

\hypertarget{ref-Bolker2009}{}
Bolker, Benjamin M, Mollie E Brooks, Connie J Clark, Shane W Geange,
John R Poulsen, M Henry H Stevens, and Jada-Simone S White. 2009.
``Generalized Linear Mixed Models: A Practical Guide for Ecology and
Evolution.'' \emph{Trends Ecol. Evol.} 24 (3): 127--35.

\hypertarget{ref-Janzen1998}{}
Frederic J Janzen and Hal S Stearn. 1998. ``Logistic Regression for
Empirical Studies of Multivariate Selection.'' \emph{Evolution} 52 (6):
1564--71.

\hypertarget{ref-Gripenberg2010}{}
Gripenberg, Sofia, Peter J Mayhew, Mark Parnell, and Tomas Roslin. 2010.
``A Meta-Analysis of Preference-Performance Relationships in
Phytophagous Insects.'' \emph{Ecol. Lett.} 13 (3): 383--93.

\hypertarget{ref-Hawkins1997}{}
Hawkins, Bradford A, Howard V Cornell, and Michael E Hochberg. 1997.
``Predators, Parasitoids, and Pathogens as Mortality Agents in
Phytophagous Insect Populations.'' \emph{Ecology} 78 (7). Ecological
Society of America: 2145--52.

\hypertarget{ref-Heath2018}{}
Heath, Jeremy J, Patrick Abbot, and John O Stireman 3rd. 2018.
``Adaptive Divergence in a Defense Symbiosis Driven from the Top down.''
\emph{Am. Nat.} 192 (1): E21--E36.

\hypertarget{ref-Singer1986}{}
Singer, Michael C. 1986. ``The Definition and Measurement of Oviposition
Preference in Plant-Feeding Insects.'' In \emph{Insect-Plant
Interactions}, edited by James R Miller and Thomas A Miller, 65--94. New
York, NY: Springer New York.

\hypertarget{ref-Weis1983}{}
Weis, Arthur E, Peter W Price, and Michael Lynch. 1983. ``Selective
Pressures on Clutch Size in the Gall Maker Asteromyia Carbonifera.''
\emph{Ecology} 64 (4). Ecological Society of America: 688--95.
